Learning Discrete Bayesian Networks from Continuous Data

### 2015-12-02 BN_discretizer_paper.tex
A partial update according to Mykel's comments

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
%\begin{filecontents*}{example.eps}
  %!PS-Adobe-3.0 EPSF-3.0
  %%BoundingBox: 19 19 221 221
  %%CreationDate: Mon Sep 29 1997
  %%Creator: programmed by hand (JK)
  %%EndComments
%  gsave
%  newpath
%    20 20 moveto
%    20 220 lineto
%    220 220 lineto
%    220 20 lineto
%  closepath
%  2 setlinewidth
%  gsave
%    .4 setgray fill
%  grestore
%  stroke
%  grestore
%\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[round]{natbib}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{tikz}
\usepgfplotslibrary{groupplots}
\usepackage{lscape}
% \usepackage{caption}
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage{BOONDOX-cal} % for lowercase caligraphic
\usepackage[font=small,skip=5pt]{caption}
\usepackage[hyphens]{url}

@@

Real data often contains a mixture of discrete and continuous variables, but many Bayesian network structure learning and inference algorithms assume all random variables are discrete.
Continuous variables are often discretized, but the choice of discretization policy has significant impact on the accuracy, speed, and interpretability of the resulting models.
This paper introduces a principled Bayesian discretization method for continuous variables in Bayesian networks with quadratic complexity instead of the cubic complexity of other standard techniques.
Empirical demonstrations show that the proposed method is superior to the state of the art.
In addition, this paper shows how to incorporate existing methods into the structure learning process to discretize all continuous variables and simultaneously learn Bayesian network structures.
The proposed method was incorporated and its performance was empirically demonstrated.

@@

There are three common approaches to extending Bayesian networks to continuous variables.
The first is to model the conditional probability density of each continuous variable using specific families of parametric distributions, and then to redesign Bayesian network learning and inference algorithms based on the parameterizations.
One successful example of parametric continuous distributions in Bayesian networks is the Gaussian graphical model \citep{Weiss_2011}.
%For other parametric distributions and non-linear functions \citep{Ihler_2009}, however, these redesigned algorithms can be computationally expensive.

@@

Results in this work suggest that the MDL method suffers from low sensitivity to discretization edge locations and returns too few discretization intervals for continuous variables.
This is caused by MDL's use of mutual information to measure the quality of discretization edges.
Mutual information, which is composed of empirical probabilities computed using event count ratios, varies less severely with the positions of discretization edges than the method we suggest in this article.

@@

MODL \citep{Boulle_2006} is a Bayesian method for discretizing a continuous feature according to a class variable, which selects the model with maximum probability given the data.
The MODL method uses dynamic programming to find the optimal discretization policy for a continuous variable given a discrete class variable, and has an $\bigo\paren{n^3 + r \cdot n^2}$ runtime, where $r$ is the number of class variable instantiations.
\cite{Lustgarten_2011} suggest several formulations for the prior term $P(\textit{Model})$, and the asymptotic equivalence between MDL and MODL on the single-variable, single-class problem was examined by \citet{VL_2000}.

@@

\begin{figure}[ht]
  \centering
  \input{graph1}
  \caption{Factorization of $P(_{-\cX} \mid \Lambda)$}
  \label{fig:example_factorization}
\end{figure}

@@

\begin{figure}[ht]
 \centering
     \input{graph2_right}
  \caption{An example network}
   \label{fig:example_networks}
\end{figure}

@@

The global minimum of Equation~\ref{eqn:MDL} can be found using using dynamic programming.
For a given Bayesian network, the first three terms only depend on the number of discretization intervals $|X|$ and the fourth term is cumulative over the intervals.
Hence, if $\Lambda = \anglebrackets{\gamma_1, \gamma_2, \ldots, \gamma_k}$ is a MDL optimal discretization policy with $k$ intervals, then $\Lambda' = \anglebrackets{\gamma_1, \gamma_2, \ldots, \gamma_{k-1}}$ is a MDL optimal discretization policy for the corresponding subproblem with $k-1$ intervals.
This procedure takes runtime $\bigo \paren{ n^3 + \paren{ n_c \cdot {L}^{n_s}  + {L}^{n_p}} \cdot n^2 }$.
All variables are defined in Section \ref{subsec:algo}.
Therefore, the runtime of the proposed method in this paper is shorter than the MDL discretization method by $\bigo \paren{n^3}$, since the former has a more efficient form of dynamical programming.

@@

\begin{figure}[ht]
  \centering
   \input{graph_autocar_1}
   \caption{Bayesian network structure obtained from running K2 on the prediscretized Auto MPG dataset.}
  \label{fig:auto_graph_1}
\end{figure}

@@

\begin{table}[ht]
  \centering
  \caption{
    Discretization result of the Auto MPG dataset with fixed structure from Figure~\ref{fig:auto_graph_1}.
    The first five rows list the discretization edges and the last row lists the mean cross-validated log-likelihood; positive values are better.
    %The MDL method does not produce any discretization edges.
  }
  \input{table_autoMPG_1.tex}
  \label{table:auto_disc_table_1}
\end{table}

@@

\begin{figure}[ht]
  \centering
  \input{graph_auto_exp1_distr_1_3}
  \caption{
    Comparison of the Bayesian discretization policy for variables $A$ and $C$ to the original Auto MPG data learned with a fixed network.
    The marginal probability density closely matches the data.
  }
  \label{fig:auto_exp1_distr_1_3}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \input{graph_autocar_2}
  \caption{The discrete-valued Bayesian network learned from the Auto MPG dataset using the Bayesian method.}
  \label{fig:auto_graph_2}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \input{graph_auto_exp2_distr_1_3}
  \caption{
    Comparison of the Bayesian discretization policy for variables $A$ and $C$ to the original Auto MPG data learned simultaneously with the network structure.
    Although the number of discretization edges is less than those in Figure~\ref{fig:auto_exp1_distr_1_3}, the probability distribution still closely matches the original data.
  }
  \label{fig:auto_exp2_distr_1_3}
\end{figure}

@@

Some discretization edges appear in all three discretization methods, such as \num{1.42} and \num{2.35} for variable $C$ and \num{0.785} for variable $L$.
This indicates that MDL indeed can find some important discretization edges, but it is not sensitive enough to find more edges.

\begin{figure}[ht]
  \centering
  \scalebox{0.8}{\input{graph_wine_1}}
  \caption{Bayesian network structure obtained from running K2 on the prediscretized Wine dataset}
  \label{fig:wine_graph_1}
\end{figure}

\begin{table}
  \centering
  \caption{
    Results from discretization of the Wine MPG dataset with fixed structure.
    % The first thirteen rows list the discretization edges for each continuous variable.
    % The last row lists the mean cross-validated log-likelihood; positive values are better.
    % The optimal Bayesian approach has the best discretization result in terms of likelihood, but the heuristic method obtains a similar result with nearly the same likelihood.
    % The MDL method produces discretization edges for more than half of the continuous variables but does not produce enough discretization intervals.
    % Some discretization edges appear in the results of all three methods.
    Bold discretization edges were identified by both methods.
  }
  \scalebox{1.0}{
  \input{table_wine_1.tex}
  }
  \label{table:wine_disc_table_1}
\end{table}

@@

\begin{figure}[ht]
  \centering
  \input{graph_wine_exp1_distr}
  \caption{Comparison of the discretization policies for variables $A$ and $C$ obtained using the Bayesian and MDL methods}
  \label{fig:wine_exp1_distr}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \scalebox{0.7}{\input{graph_wine_2}}
  \caption{The discrete-valued Bayesian network learned from the Wine dataset using the Bayesian method.}
  \label{fig:wine_graph_2}
\end{figure}

\begin{figure}[ht]
  \centering
  \input{graph_wine_exp2_distr_5_11}
  \caption{
    The discretization policy for variables $E$ and $K$ on the learned network.
    The discretization edge $E = 17.9$ also appears in the policies for the fixed network.
  }
  \label{fig:wine_exp2_distr}
\end{figure}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% Housing
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\subsection{Data 3: Housing}
\label{subsec:housing}

@@

\begin{figure}[ht]
  \centering
  \scalebox{0.8}{\input{graph_housing_1}}
  \caption{Bayesian network structure obtained from running K2 on the prediscretized Housing dataset}
  \label{fig:housing_graph_1}
\end{figure}

\begin{table}
  \centering
  \caption{
    Discretization policy summary of the Housing dataset based on the fixed network structure shown.
    The first twelve rows show the numbers of discretization intervals and the last row is the mean cross-validated log-likelihood.% based on each discretization result.
    % The heuristic method is over-sensitive and discretizes some variables into too many intervals.
    % The MDL method is less sensitive and on many cases does not produce any discretization intervals.
    % The optimal Bayesian approach produces a reasonable number of discretization edges.
  }
  \input{table_housing_1.tex}
  \label{table:housing_disc_table_1}
\end{table}

@@

\begin{figure}[ht]
  \centering
  \input{graph_housing_exp1_distr_3_5}
  \caption{
    The discretization policies for variables $C$ and $E$ learned using the Bayesian approach on the fixed network.
    The scatter points were jittered to reveal the repeated values in the dataset.
    The repeated values lead to a high number of discretization edges, because repeated values can easily form discretized regions on their own.
  }
  \label{fig:housing_exp1_distr_3_5}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \input{graph_housing_exp1_distr_8_5}
  \caption{
    The discretization policies for variables $H$ and $E$ learned using the optimal Bayesian approach on the fixed network shown in Figure~\ref{fig:housing_graph_1}.
    The values for variable $H$ are less frequently repeated, thus producing fewer discretization intervals than $E$.
  }
  \label{fig:housing_exp1_distr_8_5}
\end{figure}

\begin{figure}[H]
  \centering
  \scalebox{0.8}{\input{graph_housing_2}}
  \caption{
    The discrete-valued Bayesian network learned from the Housing dataset using the optimal Bayesian method.
  }
  \label{fig:housing_graph_2}
\end{figure}

\begin{figure}[H]
  \centering
  \input{graph_housing_exp2_distr_3_5}
  \caption{
    The discretization policy for variables $C$ and $E$ on the learned network in Figure~\ref{fig:housing_graph_2}.
    The high number of discretization edges are caused by repeated values in $C$ and $E$.
    Note that the two regions at $(C,E) = (3.9,0.65)$ and $(6.8,0.44)$ are color-saturated due to the high concentration of repeated values.
  }
  \label{fig:housing_exp2_distr_3_5}
\end{figure}

\begin{figure}[H]
  \centering
  \input{graph_housing_exp2_distr_8_5}
  \caption{
    The discretization policy for variables $H$ and $E$ on the learned network in Figure~\ref{fig:housing_graph_2}.
    The number of discretization intervals for $H$ is less than it was for the fixed network structure in Figure~\ref{fig:housing_exp1_distr_8_5}, but the discretization policy still matches the raw data.
  }
  \label{fig:housing_exp2_distr_8_5}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \input{graph_exp_discuss}
  \caption{A simple Bayesian network used to demonstrate the sensitivity of each method}
  \label{fig:exp_discuss}
\end{figure}

To qualitatively assess the sensitivity of each method to the number and position of discretization intervals, consider the Bayesian network in Figure~\ref{fig:exp_discuss}, where $\cX$ is continuous and $P_1$ and $P_2$ are discrete.
If $\Pa_{\cX} = \curly{P_1, P_2}$ and $D_{\cX} = \curly{1.0,2.0,3.0, \ldots n}$, then the corresponding objective functions for the discretization methods are:

@@

The value of the edge position term for MDL is primarily determined by the mutual information, which varies less severely than the corresponding terms in the Bayesian method.
The MDL term uses empirical probability distributions, based off of ratios of counts, whereas the Bayesian method uses factorial terms, and thus the MDL method varies less.
The MDL method is therefore less sensitive to the discretization edges.
This sensitivity gives rise to the relative performance of the two methods in the experiments conducted above.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
This paper introduced a principled discretization method for continuous variables in Bayesian networks with quadratic complexity instead of the cubic complexity of other standard techniques.
Empirical demonstrations show that the proposed method is superior to the state of the art.
In addition, this paper shows how to incorporate existing methods into the structure learning process to discretize all continuous variables and simultaneously learn Bayesian network structures.
The proposed method was incorporated and its superior performance was empirically demonstrated.
All software is publically available to support discretization research, and can be found at \url{github.com/sisl/LearnDiscreteBayesNets.jl}.

@@

The Bayesian network structure in Figure~\ref{fig:auto_graph_3} and the discretization results in Table~\ref{table:auto_disc_table_3} demonstrate that the MDL method does not always produce fewer discretization intervals than the Bayesian approach and that the MDL method can produce discretization policies with comparable likelihood.
The Bayesian network in Figure~\ref{fig:auto_graph_3} was obtained by running Algorithm~\ref{alg:structure_learn} fifty times on the Auto MPG dataset with a maximum of two parents per variable.
The network is more well-structured than the one in Figure~\ref{fig:auto_graph_1} because the discretization policies were simultaneously learned with the network structure.
Many edges appear in both discretization results, and are bolded in Table~\ref{table:auto_disc_table_3}.

\begin{figure}[ht]
  \centering
  \scalebox{0.8}{\input{graph_autocar_3}}
  \caption{A Bayesian network structure for the Auto MPG dataset that leads to comparable discretization policies for the MDL and Bayesian methods.}
  \label{fig:auto_graph_3}
\end{figure}

\begin{table}[ht]
  \centering
  \caption{
    The discretization policies for the Auto MPG dataset with the fixed structure shown in Figure~\ref{fig:auto_graph_3}.
    % The first five rows list the discretization edges for each continuous variable.
    % The last row shows the mean cross-validated log-likelihood under each policy.
    The methods have the same discretization for variable $A$, and the log-likelihood for the Bayesian method is only slightly better than that for the MDL method.
  }
\input{table_autoMPG_3.tex}
  \label{table:auto_disc_table_3}
\end{table}
\end{document}
% end of file template.tex

>>>>

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
%\begin{filecontents*}{example.eps}
  %!PS-Adobe-3.0 EPSF-3.0
  %%BoundingBox: 19 19 221 221
  %%CreationDate: Mon Sep 29 1997
  %%Creator: programmed by hand (JK)
  %%EndComments
%  gsave
%  newpath
%    20 20 moveto
%    20 220 lineto
%    220 220 lineto
%    220 20 lineto
%  closepath
%  2 setlinewidth
%  gsave
%    .4 setgray fill
%  grestore
%  stroke
%  grestore
%\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[round]{natbib}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}
\usepgfplotslibrary{groupplots}
\usepackage{lscape}
% \usepackage{caption}
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage{BOONDOX-cal} % for lowercase caligraphic
\usepackage[font=small,skip=5pt]{caption}
\usepackage[hyphens]{url}

@@

Real data often contains a mixture of discrete and continuous variables, but many Bayesian network structure learning and inference algorithms assume all random variables are discrete.
Continuous variables are often discretized, but the choice of discretization policy has significant impact on the accuracy, speed, and interpretability of the resulting models.
This paper introduces a principled Bayesian discretization method for continuous variables in Bayesian networks with quadratic complexity instead of the cubic complexity of other standard techniques.
Empirical demonstrations show that the proposed method is superior to the state of the art.
In addition, this paper shows how to incorporate existing methods into the structure learning process to discretize all continuous variables and simultaneously learn Bayesian network structures.
%The proposed method was incorporated and its performance was empirically demonstrated.

@@

There are three common approaches to extending Bayesian networks to continuous variables.
The first is to model the conditional probability density of each continuous variable using specific families of parametric distributions, and then to redesign Bayesian network learning and inference algorithms based on the parameterizations.
One example of parametric continuous distributions in Bayesian networks is the Gaussian graphical model \citep{Weiss_2011}.
%For other parametric distributions and non-linear functions \citep{Ihler_2009}, however, these redesigned algorithms can be computationally expensive.

@@

Results in this work suggest that the MDL method suffers from low sensitivity to discretization edge locations and returns too few discretization intervals for continuous variables.
This is caused by MDL's use of mutual information to measure the quality of discretization edges.
Mutual information, which is composed of empirical probabilities computed using event count ratios, varies less significantly with the positions of discretization edges than the method we suggest in this article.

@@

MODL \citep{Boulle_2006} is a Bayesian method for discretizing a continuous feature according to a class variable, which selects the model with maximum probability given the data.
The MODL method uses dynamic programming to find the optimal discretization policy for a continuous variable given a discrete class variable, and has an $\bigo\paren{n^3 + r \cdot n^2}$ runtime, where $r$ is the number of class variable instantiations.
\cite{Lustgarten_2011} suggest several formulations for the prior over models. The asymptotic equivalence between MDL and MODL on the single-variable, single-class problem was examined by \citet{VL_2000}.

@@

\begin{figure}[ht]
  \centering
  \input{Fig1}
  \caption{Factorization of $P(_{-\cX} \mid \Lambda)$.}
  \label{fig:example_factorization}
\end{figure}

@@

\begin{figure}[ht]
 \centering
     \input{Fig2}
  \caption{An example network.}
   \label{fig:example_networks}
\end{figure}

@@

The global minimum of Equation~\ref{eqn:MDL} can be found using dynamic programming.
For a given Bayesian network, the first three terms only depend on the number of discretization intervals $|X|$ and the fourth term is cumulative over the intervals.
Hence, if $\Lambda = \anglebrackets{\gamma_1, \gamma_2, \ldots, \gamma_k}$ is a MDL optimal discretization policy with $k$ intervals, then $\Lambda' = \anglebrackets{\gamma_1, \gamma_2, \ldots, \gamma_{k-1}}$ is a MDL optimal discretization policy for the corresponding subproblem with $k-1$ intervals.
This procedure takes runtime $\bigo \paren{ n^3 + \paren{ n_c \cdot {L}^{n_s}  + {L}^{n_p}} \cdot n^2 }$.
All variables are defined in Section \ref{subsec:algo}.
Therefore, the runtime of the proposed method in this paper is shorter than the MDL discretization method by $\bigo \paren{n^3}$, since the former has a more efficient form of dynamical programming.

@@

\begin{figure}[ht]
  \centering
   \input{Fig3}
   \caption{Bayesian network structure obtained from running K2 on the prediscretized Auto MPG dataset.}
  \label{fig:auto_graph_1}
\end{figure}

@@

\begin{table}[ht]
  \centering
  \caption{
    Discretization result of the Auto MPG dataset with fixed structure from Figure~\ref{fig:auto_graph_1}.
    The first five rows list the discretization edges and the last row lists the mean cross-validated log-likelihood; positive values are better.
    %The MDL method does not produce any discretization edges.
  }
  \input{Table1.tex}
  \label{table:auto_disc_table_1}
\end{table}

@@

\begin{figure}[ht]
  \centering
  \input{Fig4}
  \caption{
    Comparison of the Bayesian discretization policy for variables $A$ and $C$ to the original Auto MPG data learned with a fixed network.
    The marginal probability density closely matches the data.
  }
  \label{fig:auto_exp1_distr_1_3}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \input{Fig5}
  \caption{The discrete-valued Bayesian network learned from the Auto MPG dataset using the Bayesian method.}
  \label{fig:auto_graph_2}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \input{Fig6}
  \caption{
    Comparison of the Bayesian discretization policy for variables $A$ and $C$ to the original Auto MPG data learned simultaneously with the network structure.
    Although the number of discretization edges is less than those in Figure~\ref{fig:auto_exp1_distr_1_3}, the probability distribution still closely matches the original data.
  }
  \label{fig:auto_exp2_distr_1_3}
\end{figure}

@@

Some discretization edges appear in all three discretization methods, such as \num{1.42} and \num{2.35} for variable $C$ and \num{0.785} for variable $L$.
This indicates sufficiently MDL indeed can find some important discretization edges, but it is not sensitive to find more edges.

\begin{figure}[ht]
  \centering
  \scalebox{0.8}{\input{Fig7}}
  \caption{Bayesian network structure obtained from running K2 on the prediscretized Wine dataset.}
  \label{fig:wine_graph_1}
\end{figure}

\begin{table}
  \centering
  \caption{
    Results from discretization of the Wine MPG dataset with fixed structure.
    % The first thirteen rows list the discretization edges for each continuous variable.
    % The last row lists the mean cross-validated log-likelihood; positive values are better.
    % The optimal Bayesian approach has the best discretization result in terms of likelihood, but the heuristic method obtains a similar result with nearly the same likelihood.
    % The MDL method produces discretization edges for more than half of the continuous variables but does not produce enough discretization intervals.
    % Some discretization edges appear in the results of all three methods.
    Bold discretization edges were identified by both methods.
  }
  \scalebox{1.0}{
  \input{Table2.tex}
  }
  \label{table:wine_disc_table_1}
\end{table}

@@

\begin{figure}[ht]
  \centering
  \input{Fig8}
  \caption{Comparison of the discretization policies for variables $A$ and $C$ obtained using the Bayesian and MDL methods.}
  \label{fig:wine_exp1_distr}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \scalebox{0.7}{\input{Fig9}}
  \caption{The discrete-valued Bayesian network learned from the Wine dataset using the Bayesian method.}
  \label{fig:wine_graph_2}
\end{figure}

\begin{figure}[ht]
  \centering
  \input{Fig10}
  \caption{
    The discretization policy for variables $E$ and $K$ on the learned network.
    The discretization edge $E = 17.9$ also appears in the policies for the fixed network.
  }
  \label{fig:wine_exp2_distr}
\end{figure}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% Housing
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\subsection{Data 3: Housing}
\label{subsec:housing}

@@

\begin{figure}[ht]
  \centering
  \scalebox{0.8}{\input{Fig11}}
  \caption{Bayesian network structure obtained from running K2 on the prediscretized Housing dataset.}
  \label{fig:housing_graph_1}
\end{figure}

\begin{table}
  \centering
  \caption{
    Discretization policy summary of the Housing dataset based on the fixed network structure shown.
    The first twelve rows show the numbers of discretization intervals and the last row is the mean cross-validated log-likelihood.% based on each discretization result.
    % The heuristic method is over-sensitive and discretizes some variables into too many intervals.
    % The MDL method is less sensitive and on many cases does not produce any discretization intervals.
    % The optimal Bayesian approach produces a reasonable number of discretization edges.
  }
  \input{Table3.tex}
  \label{table:housing_disc_table_1}
\end{table}

@@

\begin{figure}[ht]
  \centering
  \input{Fig12}
  \caption{
    The discretization policies for variables $C$ and $E$ learned using the Bayesian approach on the fixed network.
    The scatter points were jittered to reveal the repeated values in the dataset.
    The repeated values lead to many discretization edges because repeated values can easily form \todo{discretized regions on their own.}
  }
  \label{fig:housing_exp1_distr_3_5}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \input{Fig13}
  \caption{
    The discretization policies for variables $H$ and $E$ learned using the optimal Bayesian approach on the fixed network shown in Figure~\ref{fig:housing_graph_1}.
    The values for variable $H$ are not as repeated, thus producing fewer discretization intervals than $E$.
  }
  \label{fig:housing_exp1_distr_8_5}
\end{figure}

\begin{figure}[H]
  \centering
  \scalebox{0.8}{\input{Fig14}}
  \caption{
    The discrete-valued Bayesian network learned from the Housing dataset using the optimal Bayesian method.
  }
  \label{fig:housing_graph_2}
\end{figure}

\begin{figure}[H]
  \centering
  \input{Fig15}
  \caption{
    The discretization policy for variables $C$ and $E$ on the learned network in Figure~\ref{fig:housing_graph_2}.
    The large number of discretization edges are due to repeated values in $C$ and $E$.
    The two regions at $(C,E) = (3.9,0.65)$ and $(6.8,0.44)$ are color-saturated due to the high concentration of repeated values.
  }
  \label{fig:housing_exp2_distr_3_5}
\end{figure}

\begin{figure}[H]
  \centering
  \input{Fig16}
  \caption{
    The discretization policy for variables $H$ and $E$ on the learned network in Figure~\ref{fig:housing_graph_2}.
    There are fewer discretization intervals for $H$ than for the fixed network structure in Figure~\ref{fig:housing_exp1_distr_8_5}, but the discretization result still closely models the raw data.
  }
  \label{fig:housing_exp2_distr_8_5}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \input{Fig17}
  \caption{A simple Bayesian network used to demonstrate the sensitivity of each method}
  \label{fig:exp_discuss}
\end{figure}

To qualitatively assess the sensitivity of each method to the number and position of discretization intervals, consider the Bayesian network in Figure~\ref{fig:exp_discuss}, where $\cX$ is continuous and $P_1$ and $P_2$ are discrete.
If $\Pa_{\cX} = \curly{P_1, P_2}$ and $D_{\cX} = \curly{1,2,3, \ldots n}$, then the corresponding objective functions for the discretization methods are:

@@

The value of the edge position term for MDL is primarily determined by the mutual information, which varies less severely than the corresponding terms in the Bayesian method.
The MDL term uses empirical probability distributions based off of ratios of counts, whereas the Bayesian method uses factorial terms, and thus the MDL method varies less.
The MDL method is therefore less sensitive to the discretization edges.
This sensitivity gives rise to the relative performance of the two methods in the experiments conducted above.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
This paper introduced a principled discretization method for continuous variables in Bayesian networks with quadratic complexity instead of the cubic complexity of other standard techniques.
Empirical demonstrations show that the proposed method is superior to the state of the art.
In addition, this paper shows how to incorporate existing methods into the structure learning process to discretize all continuous variables and simultaneously learn Bayesian network structures.
The proposed method was incorporated and its superior performance was empirically demonstrated.
All software is publicly available at \url{github.com/sisl/LearnDiscreteBayesNets.jl}.

@@

The Bayesian network structure in Figure~\ref{fig:auto_graph_3} and the discretization results in Table~\ref{table:auto_disc_table_3} demonstrate that the MDL method does not always produce fewer discretization intervals than the Bayesian approach and that the MDL method can produce discretization policies with comparable likelihood.
The Bayesian network in Figure~\ref{fig:auto_graph_3} was obtained by running Algorithm~\ref{alg:structure_learn} fifty times on the Auto MPG dataset with a maximum of two parents per variable.
The network is better-structured than the one in Figure~\ref{fig:auto_graph_1} because the discretization policies were simultaneously learned with the network structure.
Many edges appear in both discretization results and are bolded in Table~\ref{table:auto_disc_table_3}.

\begin{figure}[ht]
  \centering
  \scalebox{0.8}{\input{Fig18}}
  \caption{A Bayesian network structure for the Auto MPG dataset that leads to comparable discretization policies for the MDL and Bayesian methods.}
  \label{fig:auto_graph_3}
\end{figure}

\begin{table}[ht]
  \centering
  \caption{
    The discretization policies for the Auto MPG dataset with the fixed structure shown in Figure~\ref{fig:auto_graph_3}.
    % The first five rows list the discretization edges for each continuous variable.
    % The last row shows the mean cross-validated log-likelihood under each policy.
    The methods have the same discretization for variable $A$, and the log-likelihood for the Bayesian method is only slightly better than that for the MDL method.
  }
\input{Table4.tex}
  \label{table:auto_disc_table_3}
\end{table}
\end{document}
% end of file template.tex

### 2015-12-02 Fig10.tex
A partial update according to Mykel's comments

\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			enlargelimits=0,
			zmin=0.0, zmax=1.0,
			xlabel=E,
			ylabel=K,
			colorbar,
			point meta min=0, point meta max=2.5,
			colormap={}{ gray(0cm)=(1); gray(1cm)=(0);},
			colorbar style={
			        ytick={0,0.5,...,2.6},
			        ylabel={probability density $\times 10^2$},
			        y label style={at={(axis description cs:6.5,0.5)},rotate=180,anchor=north},
			        yticklabel style={
			            text width=2.5em,
			            align=right,
			            /pgf/number format/.cd,
			                fixed,
			                fixed zerofill
			        }
			    }
			]

>>>>

\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			enlargelimits=0,
			zmin=0.0, zmax=1.0,
			xlabel=E,
			ylabel=K,
			colorbar,
			point meta min=0, point meta max=2.5,
			colormap={}{ gray(0cm)=(1); gray(1cm)=(0);},
			colorbar style={
			        ytick={0,0.5,...,2.51},
			        ylabel={probability density $\times 10^2$},
			        y label style={at={(axis description cs:4.5,0.5)},rotate=180,anchor=north},
			        yticklabel style={
			            text width=2.5em,
			            align=right,
			            /pgf/number format/.cd,
			                fixed,
			                fixed zerofill
			        }
			    }
			]

### 2015-12-02 Fig12.tex
A partial update according to Mykel's comments

\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			enlargelimits=0,
			zmin=0.0, zmax=1.0,
			xlabel=C,
			ylabel=E,
			colorbar,
			point meta min=0, point meta max=1.0,
			colormap={}{ gray(0cm)=(1); gray(1cm)=(0);},
			colorbar style={
			        ytick={0,0.2,...,1.01},
			        ylabel={probability density$ \times 10^{-1}$},
			        y label style={at={(axis description cs:6.5,0.5)},rotate=180,anchor=north},
			        yticklabels={\num{0.00},\num{0.20},\num{0.40},\num{0.60},\num{0.80},$\geq 1$},
			        y label style={at={(axis description cs:6.5,0.5)},rotate=180,anchor=north},
			        yticklabel style={
			            text width=2.5em,
			            align=right,
			            /pgf/number format/.cd,
			                fixed,
			                fixed zerofill
			        }
			    }
			]

>>>>

\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			enlargelimits=0,
			zmin=0.0, zmax=1.0,
			xlabel=$C$,
			ylabel=$E$,
			colorbar,
			point meta min=0, point meta max=1.0,
			colormap={}{ gray(0cm)=(1); gray(1cm)=(0);},
			colorbar style={
			        ytick={0,0.2,...,1.01},
			        ylabel={probability density$ \times 10^{-1}$},
			        y label style={at={(axis description cs:4.5,0.5)},rotate=180,anchor=north},
			        yticklabels={\num{0.00},\num{0.20},\num{0.40},\num{0.60},\num{0.80},$\geq 1$},
			        yticklabel style={
			            text width=2.5em,
			            align=right,
			            /pgf/number format/.cd,
			                fixed,
			                fixed zerofill
			        }
			    }
			]

### 2015-12-02 Fig13.tex
A partial update according to Mykel's comments

\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			enlargelimits=0,
			zmin=0.0, zmax=1.0,
			xlabel=$H$,
			ylabel=$E$,
			colorbar,
			point meta min=0, point meta max=1.0,
			colormap={}{ gray(0cm)=(1); gray(1cm)=(0);},
			colorbar style={
			        ytick={0,0.2,...,1.01},
			        ylabel={probability density$ \times 10^{-1}$},
			        y label style={at={(axis description cs:6.5,0.5)},rotate=180,anchor=north},
			        yticklabel style={
			            text width=2.5em,
			            align=right,
			            /pgf/number format/.cd,
			                fixed,
			                fixed zerofill
			        }
			    }
			]

>>>>

\begin{tikzpicture}
		\begin{axis}[
			view={0}{90},
			enlargelimits=0,
			zmin=0.0, zmax=1.0,
			xlabel=$H$,
			ylabel=$E$,
			colorbar,
			point meta min=0, point meta max=1.0,
			colormap={}{ gray(0cm)=(1); gray(1cm)=(0);},
			colorbar style={
			        ytick={0,0.2,...,1.01},
			        ylabel={probability density$ \times 10^{-1}$},
			        y label style={at={(axis description cs:4.5,0.5)},rotate=180,anchor=north},
			        yticklabel style={
			            text width=2.5em,
			            align=right,
			            /pgf/number format/.cd,
			                fixed,
			                fixed zerofill
			        }
			    }
			]
